A Derivative Tracking Model for Wind Power Forecast Error

### 2020-07-13 LaTeX_Code/Paper/aoas-template.tex
Added all supplementary material plots (maybe I still need the xcross one)

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/four_days/377.eps}
\includegraphics[width=0.45\textwidth]{plots/four_days/417.eps}\\
%\quad\\
%\includegraphics[width=0.45\textwidth]{plots/four_days/437.eps}
%\includegraphics[width=0.45\textwidth]{plots/four_days/477.eps}
\caption{Four 24-hour segments with the wind power real production in Uruguay (orange line) recorded every ten minutes, and the hourly wind power production forecasted by provider A (blue line).}
  \label{fig:sample_data}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/LP_6.eps}
\includegraphics[width=0.45\textwidth]{plots/MP_6.eps}\\
\quad\\
\includegraphics[width=0.45\textwidth]{plots/HP_6.eps}
\includegraphics[width=0.45\textwidth]{plots/AP_6.eps}
\caption{Wind production forecast error histograms during the period April-December 2019: low-power (upper-left plot), mid-power (upper-right plot), high-power (lower-left plot), and the global range of power (lower-right plot).}
  \label{fig:data_curtailing}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/LP.eps}
\includegraphics[width=0.45\textwidth]{plots/MP.eps}\\
\quad\\
\includegraphics[width=0.45\textwidth]{plots/HP.eps}
\includegraphics[width=0.45\textwidth]{plots/AP.eps}
\caption{Wind production forecast error histograms during the period April-December 2019 after removing 24-hour segments with artificial wind curtailment: low-power (upper-left plot), mid-power (upper-right plot), high-power (lower-left plot), and the global range of power (lower-right plot).}
  \label{fig:data_after_clean}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/LP_t.eps}
\includegraphics[width=0.45\textwidth]{plots/MP_t.eps}\\
\quad\\
\includegraphics[width=0.45\textwidth]{plots/HP_t.eps}
\includegraphics[width=0.45\textwidth]{plots/AP_t.eps}
\caption{Forecast error transition histograms during the period April-December 2019 without wind power production curtailment: low-power (upper-left plot), mid-power (upper-right plot), high-power (lower-left plot), and the global range of power (lower-right plot).}
\label{fig:error_transitions}
\end{figure}

@@

\begin{figure}[H]
\centering
  \includegraphics[width=0.45\textwidth]{plots/31.eps}\includegraphics[width=0.45\textwidth]{plots/31_2.eps}
  \caption{Pointwise confidence bands fitted, for the same daily segment, through diffusion models without derivative tracking (plot on the left) and with derivative tracking (plot on the right).}
  \label{fig:derivative_tracking}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/LP_t_LP.eps}
\includegraphics[width=0.45\textwidth]{plots/MP_t_LP.eps}\\
\quad\\
\includegraphics[width=0.45\textwidth]{plots/HP_t_LP.eps}
\includegraphics[width=0.45\textwidth]{plots/AP_t_LP.eps}
\caption{Lamperti transformed forecast error transition histograms between April and December 2019 without wind power production curtailment: low-power (upper-left plot), mid-power (upper-right plot), high-power (lower-left plot), and the global range of power (lower-right plot).}
  \label{fig:LP_transitions}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.5\textwidth]{plots/LL/Log-Likelihood.eps}\includegraphics[width=0.5\textwidth]{plots/LL/Log-Likelihood_testing.eps}
\caption{{Negative log-likelihood's level sets for the training sets (plot on the left), and for the test sets (plot on the right). All optimal values are located over the curve $\theta_0\alpha=0.097$ and $\theta_0\alpha=0.089$ for the training and test sets, respectively.}}
\label{fig:neg-LL}
\end{figure}
{On Figure (\ref{fig:neg-LL}), we can see the level sets for the negative log-likelihood for both training and test sets.} The numerical values of each relevant point can be seen in Table (\ref{tab:optimal_values}). We set the optimal parameters in the $V$-space $(\theta_0^V,\alpha^V)=(1.93,0.050)$, as it is where the negative log-likelihood for the training sets reaches its minimum value.
\begin{table}[H]
\centering
\begin{tabular}{lcccccc}
\toprule
 & \multicolumn{3}{c}{Training sets} & \multicolumn{3}{c}{Test sets} \\
 \midrule
 & $\theta_0$ & $\alpha$ & $\theta_0\alpha$ & $\theta_0$ & $\alpha$ & $\theta_0\alpha$\\
 \midrule
 Initial guess & 1.54 & 0.072 & 0.111 & 1.96 & 0.053 & 0.104 \\
 fminsearch & 1.14 & 0.076 & 0.097 & 1.64 & 0.054 & 0.089 \\
 fmincon & 1.58 & 0.062 & 0.097 & 1.63 & 0.055 & 0.089 \\
 fminunc & 1.54 & 0.063 & 0.097 & 1.96 & 0.045 & 0.089 \\
 Evaluations & 1.93 & 0.050 & 0.097 & 1.59 & 0.056 & 0.089 \\
 \bottomrule
\end{tabular}
\caption{Value points from Figure \ref{fig:neg-LL}.}
\label{tab:optimal_values}
\end{table}
{We observe that all the local (possibly global) minimizers are located over the curves $\theta_0\alpha=0.097$ and $\theta_0\alpha=0.089$ for the training and test sets, respectively. This effect shows that the optimization is more sensitive to the diffusion than to the drift.}\\

@@

\begin{figure}[ht!]
\centering
\includegraphics[width=0.45\textwidth]{plots/classic_noDelta/Optimal.eps}
\includegraphics[width=0.45\textwidth]{plots/classic_noDelta/Lamperti_Optimal.eps}
\caption{Probability histograms for error transitions. Using provider A, we overlap the real transitions from the test set with the simulated ones from the $V-$space SDE. On the left, simulations use $(\theta_0^V,\alpha^V)$. On the right, simulations use $(\theta_0^Z,\alpha^Z)$.}
\label{fig:hists}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/paths/1.eps}
\includegraphics[width=0.45\textwidth]{plots/paths/2.eps}\\
%\quad\\
%\includegraphics[width=0.45\textwidth]{plots/paths/3.eps}
%\includegraphics[width=0.45\textwidth]{plots/paths/4.eps}
\caption{Two arbitrary days with five simulated wind power production paths each.}
\label{fig:simulation_paths}
\end{figure}
Once derived optimal estimates of the parameters of the complete likelihood for Model 2, we obtain empirical pointwise confidence bands for wind power production. Figure (\ref{fig:confidence_bands}) shows the empirical pointwise confidence bands for wind power production for each day of interest, assuming Model 2 specification, a given forecaster, and 5000 simulations per day.

\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{plots/bands/1.eps}
\includegraphics[width=0.45\textwidth]{plots/bands/2.eps}\\
%\quad\\
%\includegraphics[width=0.45\textwidth]{plots/bands/3.eps}
%\includegraphics[width=0.45\textwidth]{plots/bands/4.eps}
\caption{Empirical pointwise confidence bands for the wind power production using the approximate MLEs for Model 2 $(\theta_0, \alpha ,\delta)=(2.22,0.044,0.054)$. Blue line: real production.}
\label{fig:confidence_bands}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{plots/delta/contour_delta.eps}
\caption{Initial value for $\delta$ as a function of the elements of the parameter vector $\bm{\theta}$.}
\label{fig:delta}
\end{figure}

>>>>

\begin{figure}[H]
\centering
\includegraphics[width=0.485\textwidth]{plots/four_days/377.eps}\quad
\includegraphics[width=0.485\textwidth]{plots/four_days/417.eps}\\
%\quad\\
%\includegraphics[width=0.45\textwidth]{plots/four_days/437.eps}
%\includegraphics[width=0.45\textwidth]{plots/four_days/477.eps}
\caption{Four 24-hour segments with the wind power real production in Uruguay (orange line) recorded every ten minutes, and the hourly wind power production forecasted by provider A (blue line).}
  \label{fig:sample_data}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.485\textwidth]{plots/LP_6.eps}
\includegraphics[width=0.485\textwidth]{plots/MP_6.eps}\\
\quad\\
\includegraphics[width=0.485\textwidth]{plots/HP_6.eps}
\includegraphics[width=0.485\textwidth]{plots/AP_6.eps}
\caption{Wind production forecast error histograms during the period April-December 2019: low-power (upper-left plot), mid-power (upper-right plot), high-power (lower-left plot), and the global range of power (lower-right plot).}
  \label{fig:data_curtailing}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.485\textwidth]{plots/LP.eps}
\includegraphics[width=0.485\textwidth]{plots/MP.eps}\\
\quad\\
\includegraphics[width=0.485\textwidth]{plots/HP.eps}
\includegraphics[width=0.485\textwidth]{plots/AP.eps}
\caption{Wind production forecast error histograms during the period April-December 2019 after removing 24-hour segments with artificial wind curtailment: low-power (upper-left plot), mid-power (upper-right plot), high-power (lower-left plot), and the global range of power (lower-right plot).}
  \label{fig:data_after_clean}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.485\textwidth]{plots/LP_t.eps}
\includegraphics[width=0.485\textwidth]{plots/MP_t.eps}\\
\quad\\
\includegraphics[width=0.485\textwidth]{plots/HP_t.eps}
\includegraphics[width=0.485\textwidth]{plots/AP_t.eps}
\caption{Forecast error transition histograms during the period April-December 2019 without wind power production curtailment: low-power (upper-left plot), mid-power (upper-right plot), high-power (lower-left plot), and the global range of power (lower-right plot).}
\label{fig:error_transitions}
\end{figure}

@@

\begin{figure}[H]
\centering
  \includegraphics[width=0.485\textwidth]{plots/31.eps}\quad\includegraphics[width=0.485\textwidth]{plots/31_2.eps}
  \caption{Pointwise confidence bands fitted, for the same daily segment, through diffusion models without derivative tracking (plot on the left) and with derivative tracking (plot on the right).}
  \label{fig:derivative_tracking}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.485\textwidth]{plots/LP_t_LP.eps}
\includegraphics[width=0.485\textwidth]{plots/MP_t_LP.eps}\\
\quad\\
\includegraphics[width=0.485\textwidth]{plots/HP_t_LP.eps}
\includegraphics[width=0.485\textwidth]{plots/AP_t_LP.eps}
\caption{Lamperti transformed forecast error transition histograms between April and December 2019 without wind power production curtailment: low-power (upper-left plot), mid-power (upper-right plot), high-power (lower-left plot), and the global range of power (lower-right plot).}
  \label{fig:LP_transitions}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.485\textwidth]{plots/LL/Log-Likelihood.eps}\quad\includegraphics[width=0.485\textwidth]{plots/LL/Log-Likelihood_testing.eps}
\caption{{Negative log-likelihood's level sets for the training sets (plot on the left), and for the test sets (plot on the right). All optimal values are located over the curve $\theta_0\alpha=0.097$ and $\theta_0\alpha=0.089$ for the training and test sets, respectively.}}
\label{fig:neg-LL}
\end{figure}
{On Figure (\ref{fig:neg-LL}), we can see the level sets for the negative log-likelihood for both training and test sets.} The numerical values of each relevant point can be seen in Table (\ref{tab:optimal_values}). We set the optimal parameters in the $V$-space $(\theta_0^V,\alpha^V)=(1.93,0.050)$, as it is where the negative log-likelihood for the training sets reaches its minimum value.
\begin{table}[H]
\centering
\begin{tabular}{lcccccc}
\toprule
 & \multicolumn{3}{c}{Training sets} & \multicolumn{3}{c}{Test sets} \\
 \midrule
 & $\theta_0$ & $\alpha$ & $\theta_0\alpha$ & $\theta_0$ & $\alpha$ & $\theta_0\alpha$\\
 \midrule
 Initial guess & 1.54 & 0.072 & 0.111 & 1.96 & 0.053 & 0.104 \\
 fminsearch & 1.14 & 0.076 & 0.097 & 1.64 & 0.054 & 0.089 \\
 fmincon & 1.58 & 0.062 & 0.097 & 1.63 & 0.055 & 0.089 \\
 fminunc & 1.54 & 0.063 & 0.097 & 1.96 & 0.045 & 0.089 \\
 Evaluations & 1.93 & 0.050 & 0.097 & 1.59 & 0.056 & 0.089 \\
 \bottomrule
\end{tabular}
\caption{Value points from Figure \ref{fig:neg-LL}.}
\label{tab:optimal_values}
\end{table}
{We observe that all the local (possibly global) minimizers are located over the curves $\theta_0\alpha=0.097$ and $\theta_0\alpha=0.089$ for the training and test sets, respectively. This effect shows that the optimization is more sensitive to the diffusion than to the drift.}\\

@@

\begin{figure}[ht!]
\centering
\includegraphics[width=0.485\textwidth]{plots/classic_noDelta/Optimal.eps}
\includegraphics[width=0.485\textwidth]{plots/classic_noDelta/Lamperti_Optimal.eps}
\caption{Density histograms for error transitions. Using provider A, we overlap the real transitions from the test set with the simulated ones from the $V-$space SDE. On the left, simulations use $(\theta_0^V,\alpha^V)$. On the right, simulations use $(\theta_0^Z,\alpha^Z)$.}
\label{fig:hists}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.485\textwidth]{plots/paths/1.eps}\quad
\includegraphics[width=0.485\textwidth]{plots/paths/2.eps}\\
%\quad\\
%\includegraphics[width=0.45\textwidth]{plots/paths/3.eps}
%\includegraphics[width=0.45\textwidth]{plots/paths/4.eps}
\caption{Two arbitrary days with five simulated wind power production paths each.}
\label{fig:simulation_paths}
\end{figure}
Once derived optimal estimates of the parameters of the complete likelihood for Model 2, we obtain empirical pointwise confidence bands for wind power production. Figure (\ref{fig:confidence_bands}) shows the empirical pointwise confidence bands for wind power production for each day of interest, assuming Model 2 specification, a given forecaster, and 5000 simulations per day.

\begin{figure}[H]
\centering
\includegraphics[width=0.485\textwidth]{plots/bands/1.eps}\quad
\includegraphics[width=0.485\textwidth]{plots/bands/2.eps}\\
%\quad\\
%\includegraphics[width=0.45\textwidth]{plots/bands/3.eps}
%\includegraphics[width=0.45\textwidth]{plots/bands/4.eps}
\caption{Empirical pointwise confidence bands for the wind power production using the approximate MLEs for Model 2 $(\theta_0, \alpha ,\delta)=(2.22,0.044,0.054)$. Blue line: real production.}
\label{fig:confidence_bands}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.6\textwidth]{plots/delta/contour_delta.eps}
\caption{Initial value for $\delta$ as a function of the elements of the parameter vector $\bm{\theta}$.}
\label{fig:delta}
\end{figure}
